\documentclass[11pt]{article}

\usepackage[letterpaper, margin=1in]{geometry}
\usepackage[default]{opensans}
\usepackage[skip=5pt]{parskip}
\usepackage{setspace}
\usepackage{enumitem}

\title{BRICK Bylaws}

\newcounter{h1}
\newcounter{h2}[h1]
\newcounter{h3}[h2]

\begin{document}
\newcommand{\tab}[1]{\stepcounter{h1}\par\large\textbf{\arabic{h1}. #1}\normalsize\par}
\newcommand{\subtab}[1]{\stepcounter{h2}\par\textbf{\arabic{h1}.\arabic{h2}. #1}\par}
\newcommand{\subsubtab}[1]{\stepcounter{h3}\par\small\textbf{\arabic{h1}.\arabic{h2}.\arabic{h3}. #1}\normalsize\par}
\renewcommand{\labelenumi}{\arabic{enumi}.} 
\renewcommand{\labelenumii}{\arabic{enumi}.\arabic{enumii}.}
\renewcommand{\labelenumiii}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}.}
\renewcommand{\labelenumiv}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}.\arabic{enumiv}.}

\setstretch{1.375}
\setenumerate{noitemsep}
\raggedright


\begin{center}
\Large
\textbf{BRICK Bylaws}

\normalsize
Revision 1.0 (September 29, 2025)
\end{center}


\tab{Name and purpose}

\subtab{Name}

This organization shall be known as BRICK or Brick.

\subtab{Mission statement}

BRICK is a student organization with the objective of mobilizing CMU's community toward activism centered on queer issues, by providing students with support, resources, and community. Our goal is to secure the rights and safety of queer individuals and oppose oppression in all forms.

\subtab{Purpose}

\begin{enumerate}
    \item The purpose of BRICK shall be:
          \begin{enumerate}
            \item To organize, support, and provide resources on protests or other activities which support any issue of importance to CMU's queer community.
            \item To provide support and mutual aid within CMU's queer community and to its allies.
            \item To foster a campus community inclusive of queer and marginalized groups.
          \end{enumerate}
    \item BRICK shall not operate for profit and no income of the organization may be distributed to private individuals.
    \item All activities and functions of the organization must be legal under University, local, state, and federal laws.
\end{enumerate}


\tab{Definitions}

\begin{enumerate}
  \item \textbf{Academic semester:} An academic semester shall be defined as a fall or spring semester.
  \item \textbf{Acclamation:} A vote by acclamation shall be defined as a procedure which decides a question by unanimous consent of the attendees of a COM.
  \item \textbf{Ballot item:} A ballot item shall be defined as a question to be decided by a formal vote.
  \item \textbf{Collective organization:} The collective organization shall be defined as the organization's general membership, in contrast to any of its specific committees or interest groups.
  \item \textbf{Collective organization meeting:} A collective organization meeting, or COM, shall be defined as a meeting which all members may attend as a right of membership. The primary purpose of COMs is to make decisions requiring the consensus of the organization at large. This definition is duplicated in section 5.2 ("collective organization meetings").
  \item \textbf{Committee:} A committee shall be defined as a group of members which organize actions on behalf of the organization for some specific purpose. This definition is duplicated in section 6.1 ("structure of committees").
  \item \textbf{Election:} An election shall be defined as a procedure consisting of a nomination period and a formal vote, which shall be used to select officers.
        \begin{enumerate}
          \item \textbf{General election:} A general election shall be defined as a semesterly election which shall fill all officer positions.
          \item \textbf{Special election:} A special election shall be defined as an irregularly scheduled election which shall fill one or more vacant officer positions.
        \end{enumerate}
  \item \textbf{Formal vote:} A formal vote shall be defined as an anonymous process to decide one or more questions through an in-person election. Formal votes shall occur following COMs as required, as well as for general and special elections of officers.
  \item \textbf{Interest group:} An interest group, or IG, shall be defined as a group of individuals and one or more committees with a shared focus, with the purpose of sharing information and updates relating to that focus, rather than carrying out actions, which is the duty of committees. This definition is duplicated in section 7.1 ("interest groups").
  \item \textbf{Majority:} A majority shall be defined as greater than one-half, discounting abstentions.
  \item \textbf{Member:} A member shall be defined as an individual associated with the organization who fulfills all requirements of membership and receives all rights associated with such status.
        \begin{enumerate}
          \item \textbf{Non-voting member:} A non-voting member shall be defined as a member who is not a voting member.
          \item \textbf{Voting member:} A voting member shall be defined as a member with the right to vote in COMs and formal votes.
        \end{enumerate}
  \item \textbf{Officer:} An officer shall be defined as any individual holding one or more officer positions.
  \item \textbf{Officer board:} The officer board shall be defined as the group composed of all officers of the organization.
  \item \textbf{Officer position:} An officer position shall be defined as an elected role held by one or more individuals which carries specific associated rights and duties.
  \item \textbf{Open committee meeting:} An open committee meeting shall be defined as a meeting at which a committee is present and which all members may attend as a right of membership. This definition is duplicated in section 6.4 ("open committee meetings"). 
  \item \textbf{Prohibition from membership:} Prohibition from membership shall be defined as synonymous with suspension.
  \item \textbf{Referendum:} A referendum shall be defined as a ballot item in a formal vote which does not create a resolution, but may be used as a condition of another resolution or to seek the input of the collective organization on some matter.
  \item \textbf{Resolution:} A resolution shall be defined as a formally passed decision of the collective organization, which mandates the organization to take a specific action or set of actions.
        \begin{enumerate}
          \item \textbf{Simple resolution:} A simple resolution or majority resolution shall be defined as a resolution which may be passed by acclamation or by a majority formal vote ballot item.
          \item \textbf{Supermajority resolution:} A supermajority resolution shall be defined as a resolution which may be passed only by a supermajority formal vote ballot item.
        \end{enumerate}
  \item \textbf{Single Transferable Vote (STV):} Single Transferable Vote, or STV, shall be defined as a vote counting method for ranked choice ballot items for multiple seats, which shall be counted using Meek's method.
  \item \textbf{Special interest group:} A special interest group, or SIG, shall be defined as an interest group with a limited lifespan. This definition is duplicated in section 7.2 ("special interest groups").
  \item \textbf{Supermajority:} A supermajority shall be defined as greater than two-thirds, discounting abstentions.
  \item \textbf{Suspension:} Suspension shall refer to a status applying to individuals, prohibiting membership in the organization throughout the duration of the period of their suspension, which may be indefinite. This definition is duplicated in section 3.3 ("suspension and prohibition of membership").
  \item \textbf{Weak majority:} A weak majority shall be defined as greater than or equal to one-half, discounting abstentions.
  \item \textbf{Written vote:} A written vote shall be defined as a procedure to decide a question among a group of fixed size, such as a committee or officer board. This definition is duplicated in section 8.1 ("written votes").
\end{enumerate}


\tab{Membership}

\subtab{Membership requirements}

\subtab{Rights of members}

\subtab{Suspension and prohibition of membership}


\tab{Officers}

\subtab{Purpose of officers}

\subtab{Officer positions}

\subsubtab{Clerks}
\subsubtab{Finance contact point}
\subsubtab{Communications contact point}
\subsubtab{Logistics contact point}
\subsubtab{Documents contact point}
\subsubtab{Local community contact point}

\subtab{Selection of officers}

\subsubtab{General and special elections}
\subsubtab{Nominations}
\subsubtab{Formal vote}

\subtab{Vacancies}

\subtab{Removal}


\tab{Collective organization}

\subtab{Rights and duties of the collective organization}

\subtab{Collective organization meetings}

\subtab{Resolutions}


\tab{Committees}

\subtab{Structure of committees}

\subtab{Committee formation}

\subtab{Changes to existing committees}

\subtab{Open committee meetings}


\tab{Interest groups}

\subtab{Interest groups}

\subtab{Special interest groups}


\tab{Voting}

\subtab{Written votes}

\subtab{Formal votes}

\subsubtab{Purpose of formal votes}
\subsubtab{Procedure}


\tab{Amendments}


\tab{Dissolution}

\subtab{Voluntary dissolution}

\subtab{Dissolution process}


\tab{History}
\end{document}